Categorical Foundations of E-Graphs With Binding

### 2025-07-11 POPL2026/paper/sections/categorical_semantics.tex
cosmetic changes

The collection of $\catname{SLat}$-categories and $\catname{SLat}$-functors form an $\catname{SLat}$-category $\catname{SLat}--\catname{Cat}$\footnote{More generally, if we were to define $\catname{SLat}$-natural transformations, we could view $\catname{SLat}--\catname{Cat}$ as a symmetric monoidal 2-category~\cite{Kelly2022BASICCO}.}, and for any pair of $\catname{SLat}$-categories $\mathbb{C}$ and $\mathbb{D}$, the product $\catname{SLat}$-category $\mathbb{C} \times \mathbb{D}$ is defined analogously to categories.

@@

Similar to $\catname{CS}(\Sigma)$, $\catname{CS}(\Sigma)^{+}$ can be freely generated from a set of closed $\Sigma^{+}$-terms as follows.
Every closed $\Sigma$-term is also a $\Sigma^{+}$-term.
Given two closed $\Sigma^{+}$-terms $f\colon A \to B$ and $g\colon A \to B$, $f + g\colon A \to B$ is also a closed $\Sigma^{+}$-term.
Terms of $\catname{CS}(\Sigma)^{+}$ are quotiented by the laws of closed symmetric monoidal $\catname{SLat}$-categories, including the distributivity law (\autoref{law:distributivity}).

>>>>

The collection of $\catname{SLat}$-categories and $\catname{SLat}$-functors form an $\catname{SLat}$-category $\catname{SLat}\text{--}\catname{Cat}$\footnote{More generally, if we were to define $\catname{SLat}$-natural transformations, we could view $\catname{SLat}\text{--}\catname{Cat}$ as a symmetric monoidal 2-category~\cite{Kelly2022BASICCO}.}, and for any pair of $\catname{SLat}$-categories $\mathbb{C}$ and $\mathbb{D}$, the product $\catname{SLat}$-category $\mathbb{C} \times \mathbb{D}$ is defined analogously to categories.

@@

Similar to $\catname{CS}(\Sigma)$, $\catname{CS}(\Sigma)^{+}$ can be freely generated from a set of closed $\Sigma^{+}$-terms as follows.
Every closed $\Sigma$-term is also a $\Sigma^{+}$-term.
Given two closed $\Sigma^{+}$-terms $f\colon A \to B$ and $g\colon A \to B$, $f + g\colon A \to B$ is also a closed $\Sigma^{+}$-term.
Terms of $\catname{CS}(\Sigma)^{+}$ are quotiented by the laws of closed symmetric monoidal $\catname{SLat}$-categories, including the distributivity law (\autoref{law:distributivity}).
The same applies in the presence of equations for $\catname{CS}(\Sigma,\mathcal{E})^{+}$.

### 2025-07-11 POPL2026/paper/sections/combinatorial_semantics.tex
cosmetic changes

\autoref{fig:hypergraph-with-interfaces} gives an example of an MDA cospan, which is such because it \enquote{looks like} the string diagram \comonoid.
Analogously, we can see that~\autoref{fig:hypergraph} (equipped with the interfaces choosing $\{ v_1 \}$ and $\{ v_4, v_5, v_6, v_7 \}$) fails to form an MDA cospan --- it is not monogamous for many reasons: $v_1$ has out-degree 0, $v_1, v_5, v_7$ have in-degree 2, and $v_3$ has in-degree 0; additionally, it fails be be directed acyclic due to the path along $e_4$ from $v_7$ to itself\footnote{Recall that loops are not part of the syntax of string diagrams for symmetric monoidal categories in the absence of a trace --- a relaxation of monogamy to allow for this is explored by~\citet{ghica_rewriting_2023}.}.

\begin{figure}
	\begin{subfigure}[c]{0.3\linewidth}
		\centering
		\begin{tikzpicture}[
				vertex/.style={draw, circle, fill, inner sep=1.5pt},
				hyperedge/.style={draw, rectangle}
			]
			\node[vertex, label=left:{$v_4$}] (v1) at (0,0) {};
			\node[vertex, label=left:{$v_5$}] (v2) at (1,0) {};
			\node[vertex, label=left:{$v_6$}] (v3) at (2,0) {};
			\node[vertex, label=left:{$v_7$}] (v4) at (3,0) {};
			\node[vertex, label=left:{$v_2$}] (v5) at (0.5,-2) {};
			\node[vertex, label=left:{$v_3$}] (v6) at (2,-2) {};
			\node[vertex, label=left:{$v_1$}] (v7) at (2,-3) {};

			\node[hyperedge] (e1) at (0.5,-1) {$e_3$};
			\node[hyperedge] (e2) at (2,-1) {$e_4$};
			\node[hyperedge] (e3) at (0.5,-3) {$e_1$};
			\node[hyperedge] (e4) at (2,-2.5) {$e_2$};

			\draw (v1) -- (e1);
			\draw (v2) -- (e1);
			\draw (v5) -- (e1);
			\draw (v5) edge[bend right] (e3);
			\draw (v5) edge[bend left] (e3);
			\draw (v2) edge[in=130,out=-90] (e2);
			\draw (v3) -- (e2);
			\draw (v4) edge[in=50,out=-90] (e2);
			\draw (v4) edge[in=-50,out=-90] (e2);
			\draw (v6) -- (e2);
		\end{tikzpicture}
		\caption{Hypergraph with 7 vertices and 4 hyperedges, all directed bottom-to-top; source/target vertices are ordered left-to-right.}%
		\label{fig:hypergraph}
	\end{subfigure}
	\begin{subfigure}[c]{0.3\linewidth}
		\centering
		\begin{tikzpicture}[
				vertex/.style={draw, circle, fill, inner sep=1.5pt},
				hyperedge/.style={draw, rectangle}
			]
			\node[vertex, label=left:{$v_3$}] (v1) at (0,0) {};
			\node[vertex, label=left:{$v_3$}] (v1') at (0,1.5) {};
			\node[vertex, label=left:{$v_2$}] (v2) at (1,0) {};
			\node[vertex, label=left:{$v_2$}] (v2') at (1,1.5) {};
			\node[vertex, label=left:{$v_1$}] (v3) at (0.5,-2) {};
			\node[vertex, label=left:{$v_1$}] (v3') at (0.5,-3.5) {};

			\node[hyperedge] (e1) at (0.5,-1) {$e_1$};

			\draw (v1) -- (e1);
			\draw (v2) -- (e1);
			\draw (v3) -- (e1);

			\node[draw, rounded corners, fit={(v1'.west) (v2'.west)}, inner sep=12pt] (S) {};
			\node[draw, rounded corners, fit={(v3'.west)}, inner sep=12pt] (T) {};
			\node[draw, rounded corners, fit={(v1.west) (v2.west) (v3.west)}, inner sep=12pt] (H) {};
			\draw[commutative diagrams/hookrightarrow,shorten >=2pt,shorten <=2pt] (S) -- node[left] {$g$} (H);
			\draw[commutative diagrams/hookrightarrow,shorten >=2pt,shorten <=2pt] (T) -- node[left] {$f$} (H);
		\end{tikzpicture}
		\caption{Hypergraph with interfaces}%
		\label{fig:hypergraph-with-interfaces}
	\end{subfigure}
	\caption{Example hypergraphs}%
	\label{fig:hypergraph-examples}
\end{figure}

@@

\begin{definition}
	Over a closed symmetric monoidal signature $\Sigma = (\Sigma_{O}, \Sigma_{M})$, we define a closed e-hypergraph $\mathcal{G}$ as a tuple $(V,E,s,t, l_{V}, l_{E} \textcolor{e-color}{<}, \textcolor{closed-color}{<},\consistency)$ where
	\begin{itemize}
		\item $V$ is a set of vertices.
		\item $E = \colorbox{yellow}{E} \cup \textcolor{e-color}{E} \cup \textcolor{closed-color}{E}$ is a set of hyperedges that is formed of three disjoint sets of hyperedges defined as follows.
		\item $s,t\colon E \to V^{*}$ are source and target functions.
		\item $l_{E}\colon E \to \Sigma_{M} \sqcup 1$, $l_{V}\colon V \to \Sigma_{O} \cup \text{obj}_{\multimap}$ are label functions, where $\text{obj}_{\multimap}$ consists of objects $A \multimap B$ for $A,B \in obj_{\Sigma_{O}}$.
		\item $\textcolor{e-color}{<} \subseteq \textcolor{e-color}{E} \times (V \sqcup E)$ and $\textcolor{closed-color}{<} \subseteq \textcolor{closed-color}{E} \times (V \sqcup E)$ with $\textcolor{e-color}{E}, \textcolor{closed-color}{E} \subset E$ are disjoint partial orders that are defined as a transitive closure of corresponding immediate predecessor relations that we denote as $\textcolor{e-color}{<^{\mu}}$ and $\textcolor{closed-color}{<^{\mu}}$.
		      For $x \in V \sqcup E$, the set of predecessors of $x$ is denoted as $[x) \coloneq \{x^\prime ~|~ \exists y_{1} = x^\prime, \ldots, y_{n} = x,  y_{i} \textcolor{e-color}{<^{\mu}} y_{i + 1} \text{ or } y_{i} \textcolor{closed-color}{<^{\mu}} y_{i + 1}\; \}$.
		      We call edges $e$ such that $l(e) = \bot$ hierarchical and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ top-level.
		      Each of these relations should satisfy the following:
		      \begin{enumerate}
			      \item each set of predecessors consists exclusively of hierarchical edges;
			      \item each $x$ has at most one immediate predecessor;
			      \item edges $e$ such that there is no $x$ such that $e \textcolor{e-color}{<^{\mu}} x$ or $e \textcolor{closed-color}{<^{\mu}} x$ are labelled;
			      \item the relations are closed under connectivity, \emph{i.e.}, if $v \in s(e)$ then $e^\prime \textcolor{e-color}{<^{\mu}} e$ if and only if $e^\prime \textcolor{e-color}{<^{\mu}} v$, similarly for $v \in t(e)$ and $\textcolor{closed-color}{<^{\mu}}$.
		      \end{enumerate}
		\item $\consistency$ is a consistency relation which is given by the union of a family of equivalence relations $\consistency_p$ on each set $\{x \in V \sqcup E ~|~ p \textcolor{e-color}{<^\mu} x\}$ of elements which share the same parent where each relation is also closed under connectivity, i.e.\ if $v \in s(e)$ or $v \in t(e)$ such that $p \textcolor{e-color}{<^{\mu}}(v)$ and $p \textcolor{e-color}{<^{\mu}}(e)$ then $v \consistency_{p} e$.
		      We require that $\consistency_{p} \not = (E_{p} \sqcup V_{p}) \times (E_{p} \sqcup V_{p})$ where $V_{p} = \{ v ~ | ~ p \textcolor{e-color}{<^{\mu}} v\}$ and $E_{p} = \{ e ~ | ~ p \textcolor{e-color}{<^{\mu}} e\}$.
	\end{itemize}
	$\textcolor{e-color}{E}$ and $\textcolor{closed-color}{E}$ are then induced by the corresponding relations and consist of exclusively hierarchical edges, so $\colorbox{yellow}{E} = E \setminus \textcolor{e-color}{E} \cup \textcolor{closed-color}{E}$.
	When writing $x \textcolor{closed-color}{<} y$ (or $x \textcolor{e-color}{<} y$) we may refer to $x$ as the predecessor of $y$ and refer to $y$ as a successor of $x$.
	% we do not use this
	% Given a closed e-hypergraph $\mathcal{G}$ we can consider a corresponding \emph{underlying} hypergraph $\mathcal{G}^\prime$ by forgetting $\textcolor{e-color}{<}, \textcolor{closed-color}{<}$ and~$\consistency$.
\end{definition}

@@

Consider an example in the middle part of~\autoref{fig:e-cospan-example}.
The closed e-hypergraph in the example is given by $V = \{v_{1}, \ldots w_{10}\}$, $E = \{e_{1}, \ldots, e_{7}\}$, $\textcolor{e-color}{E} = \{e_{1}\}$, $\textcolor{closed-color}{E} = \{e_{2}\}$, $l_{E} = {e_{1} \mapsto \bot, e_{2} \mapsto \bot, e_{3} \mapsto g, \ldots, e_{7} \mapsto f}$ and then we have
\[
	\begin{array}{cccc}
		v_{4} \textcolor{e-color}{<^{\mu}} e_{1}  & v_{7} \textcolor{closed-color}{<^{\mu}} e_{2} & v_{4} \consistency e_{2}   & v_{4} \not \consistency v_{10} \\
		e_{3} \textcolor{e-color}{<^{\mu}} e_{1}  & v_{8} \textcolor{closed-color}{<^{\mu}} e_{2} & e_{2} \consistency u_{1}   & e_{2} \not \consistency e_{6}  \\
		e_{2} \textcolor{e-color}{<^{\mu}} e_{1}  & e_{5} \textcolor{closed-color}{<^{\mu}} e_{2} & e_{2} \consistency e_{4}   & e_{4} \not \consistency e_{7}  \\
		\ldots                                    & \ldots                                        & \ldots                     & \ldots                         \\
		v_{12} \textcolor{e-color}{<^{\mu}} e_{1} & w_{9} \textcolor{closed-color}{<^{\mu}} e_{2} & v_{11} \consistency v_{12} & w_{6} \not \consistency w_{10}
	\end{array}
\]

and, in particular, we have $[e_{5}) = \{e_2, e_1\}$.
$\consistency_{e_{1}}$ defines a partition on immediate successors of $e_{1}$ which we delimit by a vertical dashed line.
We depict edges in $\textcolor{e-color}{E}$ with a dashed box, and edges in $\textcolor{closed-color}{E}$ with a round box.
\begin{figure}
	\[
		\adjustbox{scale=0.6}{
			\tikzfig{./figures/closed_iso_example_2_2}
		}
	\]
	\captionsetup{skip=0pt, belowskip=-1ex}
	\caption{Cospan of closed e-hypergraphs example}
	\label{fig:e-cospan-example}
\end{figure}
\begin{definition}[Closed e-hypergraph homomorphism]
	A homomorphism $\phi\colon \mathcal{F}\to\mathcal{G}$ between closed e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V\colon  V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E\colon E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that:
	\begin{enumerate}
		\item $\phi$ is a hypergraph homomorphism;
		\item $\phi_{E}(\textcolor{e-color}{E_{\mathcal{F}}}) \subseteq \textcolor{e-color}{E_{\mathcal{G}}}$ and $\phi_{E}(\textcolor{closed-color}{E_{\mathcal{F}}}) \subseteq \textcolor{closed-color}{E_{\mathcal{G}}}$;
		\item for $v \in V_{\mathcal{F}}$ and $e \in E_{\mathcal{F}}$
		      \begin{equation*}
			      \text{if } e \textcolor{e-color}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{e-color}{<_{\mathcal{G}}^{\mu}} \phi(v),
			      \qquad
			      \text{if } e \textcolor{closed-color}{<_{\mathcal{F}}^{\mu}} v \text{ then } \phi(e) \textcolor{closed-color}{<_{\mathcal{G}}^{\mu}} \phi(v),
		      \end{equation*}
		      and for $e_1, e_2 \in E_{\mathcal{F}}$
		      \begin{equation*}
			      \text{if } e_1 \textcolor{e-color}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{e-color}{<_{\mathcal{G}}^{\mu}} \phi(e_2),
			      \qquad
			      \text{if } e_1 \textcolor{closed-color}{<_{\mathcal{F}}^{\mu}} e_2 \text{ then } \phi(e_1) \textcolor{closed-color}{<_{\mathcal{G}}^{\mu}} \phi(e_2);
		      \end{equation*}
		\item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 ~\consistency_{\mathcal{F}}~ x_2$ implies $\phi(x_1) ~\consistency_{\mathcal{G}}~ \phi(x_2)$.
	\end{enumerate}
	% If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \emph{strict}.
\end{definition}
The conditions on closed e-hypergraph homomorphisms require preserving immediate predecessors and the consistency relation.
\begin{definition}[Category of closed e-hypergraphs]
	The category of closed e-hypergraphs, denoted  $\catname{CEHyp(\Sigma)}$,  has closed e-hypergraphs as objects and closed e-hypergraph homomorphisms as morphisms.
\end{definition}

@@

\begin{definition}[Isomorphic cospans]
	Consider a relation
	\ifdefined\ONECOLUMN
		\[
			R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y) \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(y)\bigr\}
		\]
	\else
		\begin{multline*}
			R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y)\\ \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(x) \text{ and } z \textcolor{closed-color}{<^{\mu}} f_{\text{int}}(y)\bigr\}
		\end{multline*}
	\fi
	for $x, y \in n^\prime$.
	And let $S$ be its reflexive closure.
	The latter partitions $n^\prime$ into non-empty subsets $\{p_{j}\}^{k}_{j=1}$.
	We get an analogous partition for $m^\prime$.

@@

		\item for all $v,w$ in $\mathcal{G}$ in the image of $i \sqcup j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level, or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t. $v,w \textcolor{closed-color}{<^{\mu}} e$;
		\item for all $v,w$ in $\mathcal{L}^{\bot}$ in the image of $i \sqcup j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level or or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t. $v,w \textcolor{closed-color}{<^{\mu}} e$;

>>>>

An example and a non-example of an MDA cospan can be found in~\autoref{sec:appendix:hyp}.
% \autoref{fig:hypergraph-with-interfaces} gives an example of an MDA cospan, which is such because it \enquote{looks like} the string diagram \comonoid.
% Analogously, we can see that~\autoref{fig:hypergraph} (equipped with the interfaces choosing $\{ v_1 \}$ and $\{ v_4, v_5, v_6, v_7 \}$) fails to form an MDA cospan --- it is not monogamous for many reasons: $v_1$ has out-degree 0, $v_1, v_5, v_7$ have in-degree 2, and $v_3$ has in-degree 0; additionally, it fails be be directed acyclic due to the path along $e_4$ from $v_7$ to itself\footnote{Recall that loops are not part of the syntax of string diagrams for symmetric monoidal categories in the absence of a trace --- a relaxation of monogamy to allow for this is explored by~\citet{ghica_rewriting_2023}.}.

% \begin{figure}
% 	\begin{subfigure}[c]{0.3\linewidth}
% 		\centering
%         \[
%         \adjustbox{scale=0.8}{
%         \tikzfig{figures/hypergraph_example}
%         }
%         \]
% 		\caption{Hypergraph with 7 vertices and 4 hyperedges, all directed bottom-to-top; source/target vertices are ordered left-to-right.}%
% 		\label{fig:hypergraph}
% 	\end{subfigure}
% 	\begin{subfigure}[c]{0.3\linewidth}
% 		\centering
%         \[
%         \adjustbox{scale=0.8}{
%         \tikzfig{figures/cospan_hypergraph_example}
%         }
%         \]
% 		\caption{Hypergraph with interfaces}%
% 		\label{fig:hypergraph-with-interfaces}
% 	\end{subfigure}
% 	\caption{Example hypergraphs}%
% 	\label{fig:hypergraph-examples}
% \end{figure}

@@

\begin{definition}
	Over a closed symmetric monoidal signature $\Sigma = (\Sigma_{O}, \Sigma_{M})$, we define a closed e-hypergraph $\mathcal{G}$ as a tuple $(V,E,s,t, l_{V}, l_{E} \textcolor{e-color}{<}, \textcolor{closed-color}{<},\consistency)$ where
	\begin{itemize}
		\item $V$ is a set of vertices.
		\item $E = \colorbox{yellow}{E} \cup \textcolor{e-color}{E} \cup \textcolor{closed-color}{E}$ is a set of hyperedges that is formed of three disjoint sets of hyperedges defined as follows.
		\item $s,t\colon E \to V^{*}$ are source and target functions.
		\item $l_{E}\colon E \to \Sigma_{M} \sqcup 1$, $l_{V}\colon V \to \Sigma_{O} \cup \text{obj}_{\multimap}$ are label functions, where $\text{obj}_{\multimap}$ consists of objects $A \multimap B$ for $A,B \in obj_{\Sigma_{O}}$.
		\item $\textcolor{e-color}{<} \subseteq \textcolor{e-color}{E} \times (V \sqcup E)$ and $\textcolor{closed-color}{<} \subseteq \textcolor{closed-color}{E} \times (V \sqcup E)$ with $\textcolor{e-color}{E}, \textcolor{closed-color}{E} \subset E$ are disjoint partial orders that are defined as a transitive closure of corresponding immediate predecessor relations that we denote as $\textcolor{e-color}{\;<^{\mu}\;}$ and $\textcolor{closed-color}{\;<^{\mu}\;}$.
		      For $x \in V \sqcup E$, the set of predecessors of $x$ is denoted as $[x) \coloneq \{x^\prime ~|~ \exists y_{1} = x^\prime, \ldots, y_{n} = x,  y_{i} \textcolor{e-color}{\;<^{\mu}\;} y_{i + 1} \text{ or } y_{i} \textcolor{closed-color}{\;<^{\mu}\;} y_{i + 1}\; \}$.
		      We call edges $e$ such that $l(e) = \bot$ hierarchical and edges $e$ and vertices $v$ such that $[e) = \varnothing$ and $[v) = \varnothing$ top-level.
		      Each of these relations should satisfy the following:
		      \begin{enumerate}
			      \item each set of predecessors consists exclusively of hierarchical edges;
			      \item each $x$ has at most one immediate predecessor;
			      \item edges $e$ such that there is no $x$ such that $e \textcolor{e-color}{\;<^{\mu}\;} x$ or $e \textcolor{closed-color}{\;<^{\mu}\;} x$ are labelled;
			      \item the relations are closed under connectivity, \emph{i.e.}, if $v \in s(e)$ then $e^\prime \textcolor{e-color}{\;<^{\mu}\;} e$ if and only if $e^\prime \textcolor{e-color}{\;<^{\mu}\;} v$, similarly for $v \in t(e)$ and $\textcolor{closed-color}{\;<^{\mu}\;}$.
		      \end{enumerate}
		\item $\consistency$ is a consistency relation which is given by the union of a family of equivalence relations $\consistency_p$ on each set $\{x \in V \sqcup E ~|~ p \textcolor{e-color}{<^\mu} x\}$ of elements which share the same parent where each relation is also closed under connectivity, i.e.\ if $v \in s(e)$ or $v \in t(e)$ such that $p \textcolor{e-color}{\;<^{\mu}\;}(v)$ and $p \textcolor{e-color}{\;<^{\mu}\;}(e)$ then $v \consistency_{p} e$.
		      We require that $\consistency_{p} \not = (E_{p} \sqcup V_{p}) \times (E_{p} \sqcup V_{p})$ where $V_{p} = \{ v ~ | ~ p \textcolor{e-color}{\;<^{\mu}\;} v\}$ and $E_{p} = \{ e ~ | ~ p \textcolor{e-color}{\;<^{\mu}\;} e\}$.
	\end{itemize}
	$\textcolor{e-color}{E}$ and $\textcolor{closed-color}{E}$ are then induced by the corresponding relations and consist of exclusively hierarchical edges, so $\colorbox{yellow}{E} = E \setminus \textcolor{e-color}{E} \cup \textcolor{closed-color}{E}$.
	When writing $x \textcolor{closed-color}{\;<\;} y$ (or $x \textcolor{e-color}{\;<\;} y$) we may refer to $x$ as the predecessor of $y$ and refer to $y$ as a successor of $x$.
	% we do not use this
	% Given a closed e-hypergraph $\mathcal{G}$ we can consider a corresponding \emph{underlying} hypergraph $\mathcal{G}^\prime$ by forgetting $\textcolor{e-color}{<}, \textcolor{closed-color}{<}$ and~$\consistency$.
\end{definition}

@@

Consider an example in the middle part of~\autoref{fig:e-cospan-example}.
The closed e-hypergraph in the example is given by $V = \{v_{1}, \ldots w_{10}\}$, $E = \{e_{1}, \ldots, e_{7}\}$, $\textcolor{e-color}{E} = \{e_{1}\}$, $\textcolor{closed-color}{E} = \{e_{2}\}$, $l_{E} = {e_{1} \mapsto \bot, e_{2} \mapsto \bot, e_{3} \mapsto g, \ldots, e_{7} \mapsto f}$ and then we have
\[
	\begin{array}{cccc}
		v_{4} \textcolor{e-color}{\;<^{\mu}\;} e_{1}  & v_{7} \textcolor{closed-color}{\;<^{\mu}\;} e_{2} & v_{4} \consistency e_{2}   & v_{4} \not \consistency v_{10} \\
		e_{3} \textcolor{e-color}{\;<^{\mu}\;} e_{1}  & v_{8} \textcolor{closed-color}{\;<^{\mu}\;} e_{2} & e_{2} \consistency u_{1}   & e_{2} \not \consistency e_{6}  \\
		e_{2} \textcolor{e-color}{\;<^{\mu}\;} e_{1}  & e_{5} \textcolor{closed-color}{\;<^{\mu}\;} e_{2} & e_{2} \consistency e_{4}   & e_{4} \not \consistency e_{7}  \\
		\ldots                                    & \ldots                                        & \ldots                     & \ldots                         \\
		v_{12} \textcolor{e-color}{\;<^{\mu}\;} e_{1} & w_{9} \textcolor{closed-color}{\;<^{\mu}\;} e_{2} & v_{11} \consistency v_{12} & w_{6} \not \consistency w_{10}
	\end{array}
\]

and, in particular, we have $[e_{5}) = \{e_2, e_1\}$.
$\consistency_{e_{1}}$ defines a partition on immediate successors of $e_{1}$ which we delimit by a vertical dashed line.
We depict edges in $\textcolor{e-color}{E}$ with a dashed box, and edges in $\textcolor{closed-color}{E}$ with a round box.
\begin{figure}
	\[
		\adjustbox{scale=0.6}{
			\tikzfig{./figures/closed_iso_example_2_2}
		}
	\]
	\captionsetup{skip=0pt, belowskip=-1ex}
	\caption{Cospan of closed e-hypergraphs example}
	\label{fig:e-cospan-example}
\end{figure}
\begin{definition}[Closed e-hypergraph homomorphism]
	A homomorphism $\phi\colon \mathcal{F}\to\mathcal{G}$ between closed e-hypergraphs $\mathcal{F},\mathcal{G}$ is a pair of functions $\phi_V\colon  V_{\mathcal{F}} \to V_{\mathcal{G}}, \phi_E\colon E_{\mathcal{F}} \to E_{\mathcal{G}}$ such that:
	\begin{enumerate}
		\item $\phi$ is a hypergraph homomorphism;
		\item $\phi_{E}(\textcolor{e-color}{E_{\mathcal{F}}}) \subseteq \textcolor{e-color}{E_{\mathcal{G}}}$ and $\phi_{E}(\textcolor{closed-color}{E_{\mathcal{F}}}) \subseteq \textcolor{closed-color}{E_{\mathcal{G}}}$;
		\item for $v \in V_{\mathcal{F}}$ and $e \in E_{\mathcal{F}}$
		      \begin{equation*}
			      \text{if } e \textcolor{e-color}{\;<_{\mathcal{F}}^{\mu}\;} v \text{ then } \phi(e) \textcolor{e-color}{\;<_{\mathcal{G}}^{\mu}\;} \phi(v),
			      \qquad
			      \text{if } e \textcolor{closed-color}{\;<_{\mathcal{F}}^{\mu}\;} v \text{ then } \phi(e) \textcolor{closed-color}{\;<_{\mathcal{G}}^{\mu}\;} \phi(v),
		      \end{equation*}
		      and for $e_1, e_2 \in E_{\mathcal{F}}$
		      \begin{equation*}
			      \text{if } e_1 \textcolor{e-color}{\;<_{\mathcal{F}}^{\mu}\;} e_2 \text{ then } \phi(e_1) \textcolor{e-color}{\;<_{\mathcal{G}}^{\mu}\;} \phi(e_2),
			      \qquad
			      \text{if } e_1 \textcolor{closed-color}{\;<_{\mathcal{F}}^{\mu}\;} e_2 \text{ then } \phi(e_1) \textcolor{closed-color}{\;<_{\mathcal{G}}^{\mu}\;} \phi(e_2);
		      \end{equation*}
		\item for all $x_1, x_2 \in \mathcal{F}$,  $x_1 ~\consistency_{\mathcal{F}}~ x_2$ implies $\phi(x_1) ~\consistency_{\mathcal{G}}~ \phi(x_2)$.
	\end{enumerate}
	% If we further have that if $[x_1) = \varnothing$ then $[\phi(x_1)) = \varnothing$ we call the homomorphism \emph{strict}.
\end{definition}
The conditions on closed e-hypergraph homomorphisms require preserving immediate predecessors and the consistency relation.
\begin{definition}[Category of closed e-hypergraphs]
	The category of closed e-hypergraphs, denoted  $\catname{CEHyp(\Sigma)}$,  has closed e-hypergraphs as objects and closed e-hypergraph homomorphisms as morphisms.
\end{definition}

@@

\begin{definition}[Isomorphic cospans]
	Consider a relation
	\ifdefined\ONECOLUMN
		\[
			R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y) \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{closed-color}{\;<^{\mu}\;} f_{\text{int}}(x) \text{ and } z \textcolor{closed-color}{\;<^{\mu}\;} f_{\text{int}}(y)\bigr\}
		\]
	\else
		\begin{multline*}
			R = \bigl\{ x R y \text{ if } f_{\text{int}}(x) \consistency f_{\text{int}}(y)\\ \;\text{or}\; \exists z \;\text{s.t.}\; z \textcolor{closed-color}{\;<^{\mu}\;} f_{\text{int}}(x) \text{ and } z \textcolor{closed-color}{\;<^{\mu}\;} f_{\text{int}}(y)\bigr\}
		\end{multline*}
	\fi
	for $x, y \in n^\prime$.
	And let $S$ be its reflexive closure.
	The latter partitions $n^\prime$ into non-empty subsets $\{p_{j}\}^{k}_{j=1}$.
	We get an analogous partition for $m^\prime$.

@@

		\item for all $v,w$ in $\mathcal{G}$ in the image of $i \sqcup j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level, or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t. $v,w \textcolor{closed-color}{\;<^{\mu}\;} e$;
		\item for all $v,w$ in $\mathcal{L}^{\bot}$ in the image of $i \sqcup j$,  $v$ and $w$ share the same sequence of predecessors, and either $v,w$ are top-level or or $v \consistency w$, or there exists $e \in \textcolor{closed-color}{E}$ s.t. $v,w \textcolor{closed-color}{\;<^{\mu}\;} e$;
